# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r loaddataset, echo=TRUE}
#Unzip file repdata-data-activity.zip
activity <- read.csv("activity.csv", header=TRUE, na.strings="NA")

```

## What is mean total number of steps taken per day?
```{r plot, fig.height=6, fig.width=6}
library(ggplot2)

df <- aggregate(steps ~ date, data=activity, FUN=sum, na.rm=TRUE)
#plot summary of daily steps
ggplot(data=df, aes(steps)) + geom_histogram(binwidth=5000)

activity.average <- aggregate(steps ~ date, data=activity, FUN=mean, na.rm=TRUE)
activity.median <- aggregate(steps ~ date, data=activity, FUN=median, na.rm=TRUE)

activity.average
activity.median
```


## What is the average daily activity pattern?
```{r averagedailypattern, echo=TRUE}
df <- aggregate(steps ~ interval, data=activity, mean)
ggplot(df, aes(steps, interval)) + geom_line()

maxSteps <- df[which.max(df$steps), ]$interval
```
The maximum number of steps is `r maxSteps`


## Imputing missing values
```{r missingvalues, echo=TRUE}
summary(activity)
missingCount <- nrow(activity[is.na(activity$steps), ])
print(paste('The number of missing step values is ', missingCount))
```

I plan to use the weekday and Interval as the basis for imputing missing step values.
The median of step using the combination of interval and weekday (df) was created to act as a lookup.
For all values of NA in the activity data set a lookup of interval and weekday is done on the df data frame for the median values of step

```{r imputevalues, echo=TRUE}
#copy activity into new dataset
newActivity <- activity
#create new column for weekday
newActivity$weekday <- weekdays(as.Date(newActivity$date))

df <- aggregate(steps ~ interval + weekday, data=newActivity, FUN=median, na.rm=TRUE)

for(i in 1:nrow(newActivity))
{
  if(is.na(newActivity[i, 1]))
  {
    Interval <- newActivity[i, 3]
    Weekday <- newActivity[i, 4]

    # lookup values of interval and weekday in df and replace NA with the resulting step value
    newActivity[i, 1] <- df[df$interval == Interval & df$weekday == Weekday, ]$steps
  }
}

summary(newActivity)

df <- aggregate(steps ~ date + weekday, data=newActivity, FUN=sum, na.rm=TRUE)
#plot summary of daily steps with imputed values
ggplot(data=df, aes(steps)) + geom_histogram(binwidth=5000)

newActivity.average <- aggregate(steps ~ date, data=newActivity, FUN=mean, na.rm=TRUE)
newActivity.median <- aggregate(steps ~ date, data=newActivity, FUN=median, na.rm=TRUE)

newActivity.average
newActivity.median

```


## Are there differences in activity patterns between weekdays and weekends?

```{r differences, echo=TRUE}
newActivity$weekdaytype <- newActivity$weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
newActivity[newActivity$weekdaytype == TRUE, ]$weekdaytype <- 'Weekday'
newActivity[newActivity$weekdaytype == FALSE, ]$weekdaytype <- 'Weekend'

newActivity$weekdaytype <- factor(newActivity$weekdaytype)

df <- aggregate(steps ~ weekdaytype + interval, data=newActivity, mean, na.rm=TRUE)
ggplot(df, aes(steps, interval)) + geom_line() + facet_grid(weekdaytype ~ .)

```
